---
title: "Data606 Presentation 2.37"
author: "Yuen Chun Wong"
date: "September 12, 2017"
output: html_document
---


## 2.37 Portfolio Return

A portfolio's value increases by 18% during a financial boom and by 9% during normal times. It decreases by 12% during a recession. What is the expected return on this portfolio if each scenario is equally likely?

First, we create a data frame object

```{r list-data}

e <- c("Boom","Normal","Recession")
c <- c(0.333, 0.333, 0.333)
r <- c(0.18, 0.09, -0.12)
data <- data.frame(e,c,r)

names(data) <- c("Economy","Change-Pct","Return")

print(data)

```

Then, Calculate piecewise expected return.
Try to use melt() and cast() from reshape to form a "Pivot" table
The grand total of the expected returns would be the portfolio return.
```{r format-calculate-data}
data[,"Expected-Return"] <- 0.0
data["Expected-Return"] <- data["Change-Pct"] * data["Return"]

library(reshape)

data.m <- melt(data,id=c(1:3), measure=c(4))
data.c <- cast(data.m, Economy ~ variable, sum,  margins=c("grand_row"))


```

Show Result in table format.
```{r Display-Result}
library(htmlTable)

htmlTable(data.c)

```

The Portoflio return is 5%

