---
title: "Inference for numerical data"
output:
  pdf_document: default
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
---

## North Carolina births

In 2004, the state of North Carolina released a large data set containing 
information on births recorded in this state. This data set is useful to 
researchers studying the relation between habits and practices of expectant 
mothers and the birth of their children. We will work with a random sample of 
observations from this data set.

## Exploratory analysis

Load the `nc` data set into our workspace.

```{r load-data, eval=TRUE}
load("more/nc.RData")
```

We have observations on 13 different variables, some categorical and some 
numerical. The meaning of each variable is as follows.

variable         | description
---------------- | -----------
`fage`           | father's age in years.
`mage`           | mother's age in years.
`mature`         | maturity status of mother.
`weeks`          | length of pregnancy in weeks.
`premie`         | whether the birth was classified as premature (premie) or full-term.
`visits`         | number of hospital visits during pregnancy.
`marital`        | whether mother is `married` or `not married` at birth.
`gained`         | weight gained by mother during pregnancy in pounds.
`weight`         | weight of the baby at birth in pounds.
`lowbirthweight` | whether baby was classified as low birthweight (`low`) or not (`not low`).
`gender`         | gender of the baby, `female` or `male`.
`habit`          | status of the mother as a `nonsmoker` or a `smoker`.
`whitemom`       | whether mom is `white` or `not white`.

1.  What are the cases in this data set? How many cases are there in our sample?

```{r count-data-set}
 paste("There are ", nrow(nc), " cases in our sample")

```

As a first step in the analysis, we should consider summaries of the data. This
can be done using the `summary` command:

```{r summary, eval=TRUE}
summary(nc)
```

As you review the variable summaries, consider which variables are categorical 
and which are numerical. For numerical variables, are there outliers? If you 
aren't sure or want to take a closer look at the data, make a graph.

Consider the possible relationship between a mother's smoking habit and the 
weight of her baby. Plotting the data is a useful first step because it helps 
us quickly visualize trends, identify strong associations, and develop research
questions.

2.  Make a side-by-side boxplot of `habit` and `weight`. What does the plot 
highlight about the relationship between these two variables?

```{r load-libraries}

 library(ggplot2)
 library(dplyr)
 library(openintro)
 library(knitr)

```

```{r }




# Basic box plot
p <- ggplot(nc, aes(x=habit, y=weight)) + 
  geom_boxplot()
p

```

The box plots show how the medians of the two distributions compare, but we can
also compare the means of the distributions using the following function to 
split the `weight` variable into the `habit` groups, then take the mean of each
using the `mean` function.

```{r by-means, eval=TRUE}
by(nc$weight, nc$habit, mean)
```

There is an observed difference, but is this difference statistically 
significant? In order to answer this question we will conduct a hypothesis test
.

## Inference

3.  Check if the conditions necessary for inference are satisfied. Note that 
you will need to obtain sample sizes to check the conditions. You can compute 
the group size using the same `by` command above but replacing `mean` with 
`length`.

```{r by-lenth, eval=TRUE }

by(nc$weight, nc$habit, length)

```

4.  Write the hypotheses for testing if the average weights of babies born to 
smoking and non-smoking mothers are different.

 null hypothese: avg weights of babies born to smoking and non-smoking mothers are similar
 alt. hypothese: avg weights of babies born to smoking and non-smoking mothers are different
 
 Assume we are using signficant level = 0.05
 
 Check: assume there is no baby has the same mother as other babies
        the sample size is larger than 30 or 10% of population

  



```{r hypotheses-avg-weight-babies-smoking-vs-no-smoking-1}

#check the distribution
 Smoker <- nc %>% filter(habit=='smoker')
 nonSmoker <- nc %>% filter(habit=='nonsmoker')
 
 par(mfrow=c(2,1))
 hist(Smoker$weight)
 hist(nonSmoker$weight)
 
 
```



It will be a 2-side test
```{r hypotheses-avg-weight-babies-smoking-vs-no-smoking-2}


#Calculate means by habit
Summarise_by_habit <- nc %>% group_by(habit) %>% summarise(mean = mean(weight), sd = sd(weight), n=n())

kable(Summarise_by_habit)


#SE <- sqrt ( (sd1/n1) + (sd2/n2) )
#point_estimate <- (n1-n2)
#confident_intervel <- point_estmate +/1 Z*SE
#df <- min( (n1-1), (n2-1)) #to avoid Type I error
#z ~  T = (point_estimate - null (0)) / SE

#p-value on 2 - side
#p-value <- pt(z, df) *2



SE <- sqrt(
       ( subset(Summarise_by_habit, habit=='smoker')$sd^2 / subset(Summarise_by_habit, habit=='smoker')$n) + 
        ( subset(Summarise_by_habit, habit=='nonsmoker')$sd^2 / subset(Summarise_by_habit, habit=='nonsmoker')$n) 
  )

paste("Standard Error: ", SE)

#Calculate Z score
z <- ( (subset(Summarise_by_habit, habit=='smoker')$mean - subset(Summarise_by_habit, habit=='nonsmoker')$mean) - 0   )      /SE

paste("Test statistic Z: ", z)

degree_f <- min(subset(Summarise_by_habit, habit=='smoker')$n-1, subset(Summarise_by_habit, habit=='nonsmoker')$n-1 )

paste("Degree of freedom = min(n1-1, n2-1): ", degree_f)

p_value <- pt(z, degree_f) * 2

paste("P-value from t-distriubtion is ", p_value)





```

Since the P-value is very small and signifiantly lower than 0.05, we reject the null hypothese, and it is more likely average weight of babies of non-smokers and smokers are different.



Next, we introduce a new function, `inference`, that we will use for conducting
hypothesis tests and constructing confidence intervals. 

```{r inf-weight-habit-ht, eval=TRUE, tidy=FALSE}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical", order = c("smoker","nonsmoker"))
```

Let's pause for a moment to go through the arguments of this custom function. 
The first argument is `y`, which is the response variable that we are 
interested in: `nc$weight`. openre 
interested in: `"mean"` (other options are `"median"`, or `"proportion"`.) Next
we decide on the `type` of inference we want: a hypothesis test (`"ht"`) or a 
confidence interval (`"ci"`). When performing a hypothesis test, we also need 
to supply the `null` value, which in this case is `0`, since the null 
hypothesis sets the two population means equal to each other. The `alternative`
hypothesis can be `"less"`, `"greater"`, or `"twosided"`. Lastly, the `method` 
of inference can be `"theoretical"` or `"simulation"` based.

5.  Change the `type` argument to `"ci"` to construct and record a confidence 
interval for the difference between the weights of babies born to smoking and 
non-smoking mothers.

By default the function reports an interval for ($\mu_{nonsmoker} - \mu_{smoker}$)
. We can easily change this order by using the `order` argument:

```{r inf-weight-habit-ci, eval=TRUE, tidy=FALSE}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical", 
          order = c("smoker","nonsmoker"))
```

* * *

## On your own

-   Calculate a 95% confidence interval for the average length of pregnancies 
(`weeks`) and interpret it in context. Note that since you're doing inference 
on a single population parameter, there is no explanatory variable, so you can 
omit the `x` variable from the function.

```{r inf-pregnancies-ci, eval=TRUE, tidy=FALSE}
inference(y = nc$weeks, est = "mean", type = "ci", conflevel=0.95 ,null = 0, 
          alternative = "twosided", method = "theoretical" 
          )
```

-   Calculate a new confidence interval for the same parameter at the 90% 
confidence level. You can change the confidence level by adding a new argument 
to the function: `conflevel = 0.90`.

```{r inf-pregnancies-ci-90, eval=TRUE, tidy=FALSE}
inference(y = nc$weeks, est = "mean", type = "ci", conflevel=0.90 ,null = 0, 
          alternative = "twosided", method = "theoretical" 
          )
```

-   Conduct a hypothesis test evaluating whether the average weight gained by 
younger mothers is different than the average weight gained by mature mothers.


```{r inf-weight-mature-ht, eval=TRUE, tidy=FALSE}
inference(y = nc$weight, x = nc$mature, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```
 Since the P-value is larger than significant level 0.05, we can not reject the null hypothese, the average weight gained by younger mothers is similar to the average weight gained by mature mothers.



-   Now, a non-inference task: Determine the age cutoff for younger and mature 
mothers. Use a method of your choice, and explain how your method works.

 We can get the collection of the mature mothers and younger mothers and perofrm and min/max check.
 
 min(mature mother age) is slightly larger than max(younger mother ago) and it should be the cutoff line.

```{r } 
mm <- nc %>% filter(mature=='mature mom')
min(mm$mage)
ym <- nc %>% filter(mature=='younger mom')
max(ym$mage)
```

 Therefore, the cutoff line of younger mom and mature mom is age 35.

-   Pick a pair of numerical and categorical variables and come up with a 
research question evaluating the relationship between these variables. 
Formulate the question in a way that it can be answered using a hypothesis test
and/or a confidence interval. Answer your question using the `inference` 
function, report the statistical results, and also provide an explanation in 
plain language.

  Does the average of number of hospital visits during prenancy of mother with full-term babies is higher than that of mother with premie babies?
  
  Null hypotheses: avg of hospital visits during prenancy of mom with full-term babies is similar to that of mother with premie babies
  Alt. hypotheses: avg of hospital visits during prenancy of mom with full-term babies is higher that that of mother with premie babies


```{r inf-hospital-visit-premie-ht, eval=TRUE, tidy=FALSE}
inference(y = nc$visits, x = nc$premie, est = "mean", type = "ht", null = 0, 
          alternative = "greater", method = "theoretical")
```

  Since the p-value is almost zero which is similar than 0.05 significant level, we can reject the null hypothese.  It is more likely that more hospital visit would help mother to carry babies till full-term.
  

  
  
  

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0).
This lab was adapted for OpenIntro by Mine &Ccedil;etinkaya-Rundel from a lab 
written by the faculty and TAs of UCLA Statistics.
</div>
